Observation of the orientation of snow hydrometeors at sheltered and unsheltered sites

### 2024-12-01 Review_R1_answers_202412.tex

\subsection*{Editor Comment}
\textit{Thank you for submitting "Observation of falling snowflakes orientation at sheltered and unsheltered sites" [Paper \#2024GL113042] to Geophysical Research Letters. I have received 2 reviews of your manuscript, which are included below and/or attached. As you can see, the reviews indicate that major revisions are needed before we can consider proceeding with your paper. I am therefore returning the paper to you so that you can make the necessary changes.
}

\newpage
\subsection*{Reviewer 1}

\textit{The manuscript is well-written. The authors did a great job addressing the observational issues with the Multi-Angel Snowflake Camera (MASC). I particularly appreciate the reverse-engineering numerical experiments: idealized and realistic objects were generated to evaluate various methods for describing the orientation of those objects. The findings from the numerical experiment, together with the analysis of the in-situ data, will be a valuable addition to the literature and a useful resource for the remote sensing community. I recommend the manuscript be published in GRL, after the following issues are addressed:}

\subsubsection*{Major comments:}
\begin{enumerate}
    \item \textit{The discussion about the impact of the environmental winds on the orientation of the snowflakes is not sound. Wind speed alone may not be sufficient to capture this impact. The distribution of wind components likely plays a more significant role.}

    We agree with the reviewer that the distribution of wind components and turbulence at the relevant scales is probably more important than simple wind speed that we could investigate in this work (as all the MASC measurements in the database have wind speed information).

    We carefully revised the manuscript to make sure not to overstate the role of wind speed alone and also to clarify when we cannot separate between the "natural" contribution of wind on the distribution of orientations and the "artificial" contribution due to the turbulence generated by strong wind interacting with the hollow measurement system of the MASC. A few changes are highlighted in the manuscript, we can cite for example in Sec.~3:

    \blue{Atmospheric conditions are expected to play a role in shaping orientation distributions: atmospheric turbulence is probably a key driver, as well as horizontal winds interacting with the instrument frame~\cite{Fitch_AMT_2021}. }

    and later in the same section:

    \blue{This interpretation cannot unfortunately tell us any relevant information about the role of environmental turbulence, but only the likely effect of local turbulence introduced by the instrument.  Further research is therefore necessary to clarify these dynamics, in particular the role of natural turbulence in broadening the orientation distribution.}

    \hl{To beef up either the answer or the modification on the manuscript}

\end{enumerate}

\subsubsection*{Minor comments:}
\begin{enumerate}
    \item \textit{Including results on graupel in a study focused specifically on snowflakes raises questions about the sampling method.}

    Our intention was not to focus only on snowflakes (meaning aggregate snowflakes) but on all the hydrometeors that can be reasonably observed with a MASC instrument. We realized that we were not particularly careful in the naming, and we revised the manuscript with care preferring expressions as "snow", "snow particles", "snow hydrometeors", "ice-phase hydrometeors" instead of "snowflakes". The title itself is rephrased as: \blue{Observation of the orientation  of snow hydrometeors at sheltered and unsheltered sites}.

    We are however aware that retrieving orientation from MASC images of very flat hydrometeors (plates, dendrites) is likely less accurate than retrieving it from aggregate snowflakes or graupel. This happens because planar crystals become recognizable and classifiable with the method of~\cite{Praz_AMT_2017}, used in this paper, only when their basal facet face reasonably the cameras. We commented about this in the manuscript as well:

    \blue{The results, especially for planar crystal types, must be considered with extreme care. In fact, the classification method of~\cite{Praz_AMT_2017}, which is trained on a supervised set, is able to identify planar crystals when their basal facet is facing towards the MASC cameras, making them recognizable.  This introduces a bias as some orientations would not allow for classification. For example, a thin plate falling perfectly horizontal, would appear in MASC images as a thin column. }
\end{enumerate}

@@

\clearpage
\subsection*{Reviewer 2}
\textit{The authors present an interesting analysis of the orientation behavior of ice crystals and snowflakes during falling. The analysis is based on a large database of ground-based in-situ observations (Multi-Angle Snowflake Camera). Dependencies of the orientation on observational setups (wind mitigation), environmental parameters and particle properties are discussed. A simulation experiment further reveals important impacts of using different methods to estimate the average orientation angle from multiple cameras.}\\

\textit{Overall, I find the paper very well written and also the non-expert reader can easily follow. I have only a number of minor comments and suggestions. One might of course question whether the topic is of interest to a wider community, as it is a rather specific topic. However, I see the topic itself as highly relevant for active and passive microwave remote sensing but also for estimation of ice particle sedimentation velocity. Therefore, I recommend the manuscript for publication in GRL.}\\

\noindent
We thank the reviewer for their positive feedback and for the many insightful suggestions. We tried to implement most of their comments in the revised version of the manuscript, and we addressed the specific comments here below.

@@

    We now modified the manuscript to give some space to commenting and elaborating on the findings of~\cite{Stout_ACP_2024}. At first, we include this study in Table~1. Then, in the Introduction, we mention this study as:\\
    \blue{ A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$. }\\

    And in the Conclusions:\\
    \blue{It must be noted that, at least in quiescent conditions and for some types of planar crystals, spiralling falling modes around a non-zero angle have been observed in the laboratory~\cite{Stout_ACP_2024} but it remains to be seen if this behavior can be observed and isolated also in natural conditions. With this respect, our database is not refined enough to be able to isolate efficiently individual types of planar crystals. }\\

@@

    \item \textit{L. 37: Not sure if "standard deviation" is a term that is suited for the plain language summary. Maybe one can formulate simpler: "typical variability of the orientation angle\dots"}\\

    \noindent
    We agree with the reviewer, and we implemented their suggestion. \\

    \item \textit{Introduction: Honestly, comparing it with the other sections, I find the introduction a bit weak. Considering the wider audience addressed by GRL, I think the importance of ice particle orientation should be emphasized and explained more. It is certainly true that most radar polarimetric quantities are sensitive to orientation (L. 50-52). But this is by far not all in my opinion. Particle orientation also impacts the scattering signal in passive microwave (MW) observations (both from space and from ground). Many space-borne MW sensors are also polarization-sensitive. Considering the enormous efforts undertaken for example by ECMWF and other services to assimilate those all-sky MW observations, this aspect should be mentioned and cited. Another aspect is that an ice particle falling with a preferred orientation will have a very different sedimentation velocity. So characterizing the particle orientation is quite key to simulate its terminal velocity (which is quite an important parameter in any cloud model). Ice particle orientation is also relevant for determining cloud radiative properties such as optical depth or albedo.}

    We thank the reviewer for the suggestion. While we left the second part of the Introduction a bit more "specialized" (targeting an audience interested in snow imagers and their data processing), we realized that we did not provide a satasfactory broad overview, neglecting all the implications listed by the reviewer. We rephrased therefore the first part of the Introduction, removed a few technical paragraphs and included most of the suggestions of the reviewer. Some new paragraphs now read:

    \blue{The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite{Eyre_QJRMS_2022}. For example, recent studies showed that considering hydrometeor orientation in the assimilation of high frequency radiometry may improve the  forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities as the Ice Water Path~\cite{Kaur_RS_2022}.}

    \blue{The orientation and falling regime also affects the settling velocity of ice phase hydrometers, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.}\\

    \hl{I think the reviewer is right and we could make the Introduction even more broad and accessible. I did the changes above but suggestions welcome}

    \item \textit{L. 46: It would be good to add a reference for this statement. I question a bit whether this general statement is actually true. I completely agree for plate-like crystals or large aggregates. But what is for example about bullet rosettes in high cirrus clouds? Or columns/needles? Do we have solid knowledge about their falling orientation? Also the references you provide a few sentences later (\cite{Noel_JAMC_2005,Tinklenberg_JFM_2023}) focus only on planar crystals.}

    We believe the reviewer is right. We were biased tp think mostly in terms of larger snow particles (snow aggregates, graupel) or planar crystals, so our formulation was overstating the current knowledge. We rephrased fully this sentence as:
    \blue{Ice-phase hydrometeors are typically assumed to fall with a preferred orientation. It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments. Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes. }\\

@@

    \item \textit{L. 47: What do you mean by "relatively simple"? Personally, I find all those ice particles terribly complicated.}\\

    We agree that this sentence needed to be rephrased. It now reads:

    \blue{ It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments. Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}. }\\

@@

    \item \textit{Table 1: I like the idea of having a table summarizing previous studies. But are there no studies that looked at columns or needles?}\\

    This is a good point. We did not find studies providing explicitly quantitative values of orientation distributions for columns and needles (which would be a very interesting topic to tackle). There are a few important work that deal with the orientation of columns/needles, such as~\cite{Klett_JAS_1995, Reinking_JAMC_1997, Melnikov_JAOT_2013}, They however investigate the theoretical effect of turbulence on orientation (Klett) or the effect of varying orientation on remote sensing observations (Reinking, Melnikov) rather than exploring or making hypothesis about the "nature" of orientation of columns and needles.

@@

    \noindent
    It is now corrected.\\

    \item \textit{L. 157: Why "approximately"?}

    \noindent
    The reviewer is right. The sentence now reads:\\
    \blue{With \textit{Mean\_Abs} the final estimate would be 10$^\circ$, while for \textit{Mean} it would be 3.33$^\circ$}\\

    \item \textit{L. 211-215: One general thought regarding the fact that radar studies most often "retrieve" smaller range of orientation angles: If one looks at vertically pointing radar images, one always finds the lowest few hundred meters much more turbulent than the ice and snow cloud aloft. Surface friction is certainly responsible for this effect. So I am wondering if the often observable low turbulence in many ice clouds (for example indicated by small spectrum width) could be the reason why the orientation angle distribution is more narrow? Of course, the discrepancy might also be connected to unrealistic scattering properties often assumed by polarimetric shape retrievals.}

    This would be indeed a good research question. In order to answer to it, ideally it would be needed to collect MASC data (or data with a similar imager) also in the snow clouds aloft the surface. There are significant advances in this technology (imagers carried by tethered balloons for example) but we are not there yet. Airborne imagers (on research aircrafts), for which many data exist, are not able to preserve the natural orientation of the particles in their measurements. To add to the complexity, radar data themselves are often perturbed by ground clutter, they need a few hundred meters to form their beam or have internal hardware limitations (like TR-limiters) that affect short range measurements. Intuitively we can expect that, with respect to "quieter" high level clouds, near the surface we expect a broader variability driven by turbulence generated by surface friction.

    \hl{See if we can add something}

    \item \textit{Connected to this topic but even more complicated to answer is why some polarimetric radar observations seem to indicate a certain non-zero distribution of the particle orientation: \cite{Melnikov_JAOT_2013}}

    The work of~\cite{Melnikov_JAOT_2013} presents some retrievals of "fluttering intensity" from polarimetric radar data, being this intensity the standard deviation of the  distribution of orientation angles of ice-phase hydrometeors (which they assume  0-centered in their work, with zero being horizontal, although using the interval [$0,\pi$] in their definition). A possible source of confusion is that they mention sentences as "the median $\sigma_{\Theta}$ lies in an interval from 9 to 15\dots", but it is indeed the median of the standard deviations and not the median of the orientations. This study thus agrees with the other ones presented in our Table~1 and with the "standard" assumption of preferential horizontal orientation, at least in statistical sense. We decided to included this work in Table~1 and we thank the reviewer for the suggestion.

\end{enumerate}

### 2024-12-01 Grazioli_JGRA_2024_R1.tex
Updates from Overleaf

\title{Observation of falling snowflakes orientation at sheltered and unsheltered sites}

@@

\begin{keypoints}
\item Observations of snowflake orientations show symmetrical distributions around 0$^\circ$ and standard deviations that confirm some of the assumptions made in theoretical studies. 
\item Sheltered and unsheltered sites show systematic differences in orientation distributions, influenced by wind. Instrument (local perturbations) and atmospheric effects cannot be separated.
\item Numerical simulations reveal that common estimation choices made in past studies may lead to data processing artifacts.
\end{keypoints}

\begin{abstract}
Knowledge about orientation of falling snowflakes is still poorly documented with field measurements despite its importance, for example, in the interpretation of remote sensing data. This study investigates the orientation of  snowflakes using data from a Multi-Angle Snowflake Camera (MASC). We explore the impact of different observational setups (sheltered vs. unsheltered), wind conditions, hydrometeor type, and axis ratio on the orientation distributions. Numerical simulations are used to select the best orientation estimator and to understand the reason behind contrasting results reported in past literature. We find that previously reported non-zero median orientations are likely artifacts due to averaging absolute values of orientations from the three individual cameras. Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader distributions observed at unsheltered sites and/or high wind conditions.  Observed distributions may vary significantly from those assumed in previous studies, highlighting the need for further research on snowflake orientation under varying environmental conditions.
\end{abstract}

\section*{Plain Language Summary}
Understanding the orientation of falling snowflakes is important for interpreting remote sensing data and better understand snowfall micro-physics. This study uses a 3-view camera to capture detailed images of falling snowflakes and estimate their orientations. Snowflakes generally fall with their largest dimension parallel to the ground, with typical variability of the orientation angle of about 30 degrees. We found that observations at sites exposed to the wind show broader distributions of orientations compared to those in sheltered areas. Additionally, our simulations suggest that some previous studies may have incorrectly reported average orientations because of the way data was processed. These findings help refine our understanding of snowflake behavior and show the need to better comprehend the physical mechanisms affecting the distribution of snowflake orientations.

\section{Introduction}

Ice-phase hydrometeors, such as individual crystals and aggregates, are typically assumed to fall with a preferred orientation. It is generally considered that crystals populating frozen clouds aligns preferentially in a way that maximizes the area projected in the vertical direction, as deduced by remote sensing and laboratory experiments. This was considered to be true for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}. 
This assumption is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally. The concept of orientation is thus linked to the anisotropic geometry of most hydrometeors and loses any meaning for isotropic, spherical particles. In the case of snowflakes, they are often approximated by ellipsoids (or ellipses, in case of 2D images as shown in Fig.~\ref{fig:triplet}) and the orientation is thus defined by the angle between the horizontal and the major axis of the ellipsoid.

Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values. Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors. Observational studies \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence). However, they report as well non-zero mean and median orientation values, which is contradictory with what was stated previously. As we will discuss later in this study, this discrepancy is, in reality, the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation.

The problem of observing the orientations of hydrometeors while they fall, without generating disturbance induced by the measurement itself, is not trivial. Among the ground-based instruments designed to perform this task, we may cite the two-dimensional video disdrometer (2DVD) \cite{Kruger_JAOT_2002}, the multi-angle snowflake camera (MASC) \cite{Garrett_AMT_2012}, the Precipitation Imaging Package (PIP) \cite{Pettersen_atmosphere_2020}, and the Video In Situ Snowfall Sensor (VISSS) \cite{Maahn_AMT_2024}. These relatively recent instruments differ, for example, in the number of simultaneous viewpoints (one for PIP, three for MASC, two for the others), resolution (on the order of tens of micrometers for MASC and VISSS and hundreds of micrometers for 2DVD and PIP), or instrumental structure (compact for MASC and 2DVD and open-field for PIP and VISSS). These sensors allow for direct microphysical investigations based on pictures of individual hydrometeors. As mentioned above, indirect retrievals of the behavior of large populations of particles within the sampling volume of remote-sensing instruments are also commonly employed, mostly by means of dual-polarization technologies \cite{Matrosov_JAS_2005}.

@@

% ---------------
 \begin{table}
 \caption{Non-exhaustive list of research works that provide assumptions or measurements of the distribution of orientation angles of ice phase particles. $\overline{\theta}$ indicates mean orientation, $\sigma$ is the standard deviation,  $\theta_{M}$ indicates the mode of the distribution, $\theta_{Q50}$ is the median, $MAD(\theta)$ is the median absolute deviation.}
 \label{table:orientations}
 \centering
 \begin{tabular}{l p{90mm}} 
 \hline
  Ref.   & Details \\
 \hline
   \cite{Matrosov_JAM_2001} & $\overline{\theta} = 0^\circ, \sigma = 3^\circ$ \\
   \hline
   \cite{Matrosov_JAS_2005} & $\overline{\theta} = 0^\circ, \sigma = 9^\circ \pm3^\circ$ (dendrites) \\
   \hline
    \cite{Kennedy_JAMC_2011}  &   $\overline{\theta} = 0^\circ, \sigma = 15^\circ$ (dendrites) \newline $\overline{\theta} = 0^\circ, \sigma = 30^\circ$ (aggregates)     \\
   \hline
    \cite{Ryzhkov_JAMC_2011} & $\overline{\theta} = 0^\circ, \sigma = 40^\circ$ (aggregates or graupel) \\
   \hline
   \cite{Hogan_JAMC_2012} & Assumed horizontal  \\
   \hline   
   \cite{Putnam_MWR_2017} & $\overline{\theta} = 0^\circ , \sigma = 20^\circ$ (snow) \newline 
   $\overline{\theta} = 0^\circ , \sigma \in [0,60]^\circ$ (graupel) \\
   \hline
   \cite{Bukovic_JAMC_2018} &  $\overline{\theta} = 0^\circ, \sigma = 10^\circ$ (dendrites) \newline  $\overline{\theta} = 0^\circ , \sigma = 40^\circ$ (aggregates) \\
   \hline
    \cite{Matsui_JGRA_2019}  & $\overline{\theta} = 0^\circ , \sigma = 20^\circ$ (aggregates) \newline 
   $\overline{\theta} = 20^\circ , \sigma = 42^\circ$ (graupel) \\
   \hline
   \cite{Garrett_GRL_2015}$^*$ & $\theta_{Q50} = 39^\circ, \theta_{M} = 13^\circ$ (aggregates) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 16^\circ$ (rimed particles) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 20^\circ$ (graupel) \newline
   $\theta_{Q50} = 28^\circ, \theta_{M} = 13^\circ$ (low-turbulence, all types) \newline
   $\theta_{Q50}= 40^\circ, \theta_{M} = 23^\circ$ (high-turbulence, all types) \\
   \hline
   \cite{Gergely_JGRA_2016}$^*$ &  $\theta_{Q50} \in [38,47.3]^\circ$, $MAD(\theta) \in [43,51]^\circ$ \\
   \hline
   \cite{Stout_ACP_2024}$^*$ &  $\overline{\theta} \in [0,28]^\circ$ (various planar crystals, varying Re number)\\
   \hline
   \cite{Jiang_JAS_2019}$^*$ & $\theta_{M}\approx 10^\circ$ \\
   \hline
   \cite{Fitch_AMT_2021}$^*$ & $\theta_{M} = 57^\circ$ (wind $> 5$ ms$^{-1}$) \newline
   $\theta_{M} = 28^\circ$ (wind $< 1.5$ ms$^{-1}$) \\
   \hline
   \cite{Fitch_JGR_2022}$^*$ & $\theta_{M} = 3^\circ$ 
   \newline
   $\theta_{M} = 13^\circ$ (different estimations)  \\
   \hline
   \cite{Schrom_JAS_2023}$^*$ & Beta distribution (varying parameters) \\
   \hline
    \multicolumn{2}{l}{$^{*}$Papers that present orientation statistics based on field observations.}
 \end{tabular}
 \end{table}

@@

When handling three 2D images of an individual 3D object, a primary challenge is to derive a single  estimate of orientation for each snowflake. Snowflakes are commonly modeled as ellipsoids or spheroids in previous MASC studies: the orientation of the major axis of the best-fitting ellipsoid in each 2D image is estimated, and the average of the three views is considered as a proxy for orientation \cite{Garrett_GRL_2015, Gergely_JGRA_2016, Fitch_AMT_2021, Fitch_JGR_2022}. The cited studies calculate the mean of the absolute value of the three estimates, an approach referred to here as \textit{Mean\_Abs}. Alternatively, other studies perform a 3D reconstruction of the ellipsoid that best fits all three views and take the orientation of its major axis as the estimate \cite{Jiang_JAS_2019}, a family of approaches named \textit{Ellipsoid\_Fit}. Examining Table~\ref{table:orientations}, the experimental studies using these estimators report the occurrence of non-zero modal orientation values between 10$^\circ$ and 20$^\circ$, challenging the common assumption of a distribution centered at 0$^\circ$. While the non-zero modal values reported for \textit{Ellipsoid\_Fit} in \cite{Jiang_JAS_2019} are based on a small sample of aggregate snowflakes, making evaluation difficult, we demonstrate that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach simply results from averaging absolute values. This will offer a reconciliation between the common assumptions made on snowflake orientations and the contradiction of the first studies that used the MASC, a contradiction that was also noted by~\citeA{Garrett_GRL_2015} and discussed in the Conclusions of their work.

>>>>

\title{Observation of the orientation  of snow hydrometeors at sheltered and unsheltered sites}

@@

\begin{keypoints}
\item Observations of snowflake orientations show symmetrical distributions around 0$^\circ$ and standard deviations that confirm some of the assumptions made in theoretical studies. 
\item Sheltered and unsheltered sites show systematic differences in orientation distributions. Instrument (local perturbations) and atmospheric effects cannot be separated.
\item Numerical simulations reveal that common estimation choices made in past studies may lead to data processing artifacts.
\end{keypoints}

\begin{abstract}
Knowledge about orientation of falling snow is still poorly documented with field measurements despite its importance, for example, in the interpretation of remote sensing data. This study investigates the orientation of snow hydrometeors using data from a Multi-Angle Snowflake Camera (MASC). We explore the impact of different observational setups (sheltered vs. unsheltered), wind speed, hydrometeor type, and axis ratio on the orientation distributions. Numerical simulations are used to select the best orientation estimator and to understand the reason behind contrasting results reported in past literature. We find that previously reported non-zero median orientations are likely artifacts due to averaging absolute values of orientations from the three individual cameras. Observed orientations generally follow a symmetrical distribution around 0$^\circ$, with broader distributions observed at unsheltered sites and/or high wind conditions.  Observed distributions may vary significantly from those assumed in previous studies, highlighting the need for further research on hydrometeor orientations under varying environmental conditions.
\end{abstract}

\section*{Plain Language Summary}
Understanding the orientation of falling snow particles is important for interpreting remote sensing data and better understand snowfall micro-physics. This study uses a 3-view camera to capture detailed images of falling particles and estimate their orientations. They generally fall with their largest dimension parallel to the ground, with typical variability of the orientation angle of about 30 degrees. We found that observations at sites exposed to the wind show broader distributions of orientations compared to those in sheltered areas. Additionally, our simulations suggest that some previous studies may have incorrectly reported average orientations because of the way data was processed. These findings help refine our understanding of snowfall behavior and show the need to better comprehend the physical mechanisms affecting the distribution of orientations.

\section{Introduction}
Ice-phase hydrometeors are typically assumed to fall with a preferred orientation. It is generally considered that falling snow particles populating frozen clouds align preferentially in a way that maximizes the area projected in the vertical direction, as presented in studies conducting remote sensing retrievals and laboratory experiments. Such experiments are documented for planar crystals~\cite{Noel_JAMC_2005,Matrosov_JAS_2005,Tinklenberg_JFM_2024} as well as for more complex and larger snowflakes and rimed hydrometeors~\cite{Kennedy_JAMC_2011, Ryzhkov_JAMC_2011, Koebschall_EF_2023}, while less is documented for other particle types such as needles, columns or rosettes.

The preferential orientation of falling hydrometeors is exploited by dual-polarization weather radars to retrieve information about the microphysical processes occurring in precipitation and to identify which type of hydrometeors are present and where. This is relevant for the interpretation of ground-based polarimetric radar data as well as for exploiting satellite-based observations and their (all-sky) assimilation, as done for example by the European Centre for Medium-Range Weather Forecasts ECMWF~\cite<e.g.,>{Eyre_QJRMS_2022}. For example, recent studies showed that considering hydrometeor orientation in the assimilation of high frequency radiometry may improve the  forecast in areas of cloud and precipitation~\cite{Barlakas_AMT_2021} and produce significant improvements in path-integrated quantities as the Ice Water Path~\cite{Kaur_RS_2022}.

The orientation and falling regime also affects the settling velocity of ice phase hydrometers, a key parameter linking microphysical properties to other tangible quantities as the precipitation intensity, and that remains to be better understood and documented~\cite{Heymsfield_JAS_2004}.
%Except for relatively rare situations, like storm electrification \cite{Krehbiel_MAP_1996,Pineda_JGRA_2019}, the interpretation of radar meteorology relies on the assumption that anisotropic ice particles tend to fall with their major dimension oriented horizontally.
The concept of orientation is linked to the anisotropic geometry of most snow particles and loses any meaning for isotropic, spherical particles. In the case of snowfall, the hydrometeors are often approximated by ellipsoids (or ellipses, in case of 2D images as shown in Fig.~\ref{fig:triplet}) and the orientation is thus defined by the angle between the horizontal and the major axis of the ellipsoid.

Considering a population of hydrometeors rather than individual particles, as is often the case for practical applications, the distribution of orientations is often modeled with relatively narrow variability around 0$^\circ$ (i.e., horizontally oriented) values. Looking at the overview in Table~\ref{table:orientations}, one can observe that modeling studies mostly assume a Gaussian distribution, with standard deviations varying between 3$^\circ$ \cite{Matrosov_JAM_2001} and 60$^\circ$ \cite{Putnam_MWR_2017} depending on the type of hydrometeors. Observational studies conducted in the field \cite{Garrett_GRL_2015} confirm that differences exist due to hydrometeor type and environmental conditions (wind and turbulence). They report as well non-zero mean and median orientation values, which is contradictory with what was stated previously. As we will discuss later in this study, this discrepancy can be the result of the way multi-view instrumental observations are combined to retrieve an overall estimate of orientation, at least for more complex three-dimensional snow particles. A recent laboratory study using snow crystals replicas falling in an alcoholic solution~\cite{Stout_ACP_2024}, however, shows that in quiescent conditions individual planar crystals may have a spiralling fall regime with preferential orientations that depart from 0$^\circ$. A study with similar methods, but focussing on larger snow aggregates showed instead that horizontal orientation is preferred for those hydrometeors~\cite{Koebschall_EF_2023}.

The problem of observing the orientations of hydrometeors while they fall in natural outdoor conditions, without generating disturbance induced by the measurement itself, is not trivial. Among the ground-based instruments designed to perform this task, we may cite the two-dimensional video disdrometer (2DVD) \cite{Kruger_JAOT_2002}, the multi-angle snowflake camera (MASC) \cite{Garrett_AMT_2012}, the Precipitation Imaging Package (PIP) \cite{Pettersen_atmosphere_2020}, and the Video In Situ Snowfall Sensor (VISSS) \cite{Maahn_AMT_2024}. These relatively recent instruments differ, for example, in the number of simultaneous viewpoints (one for PIP, three for MASC, two for the others), resolution (on the order of tens of micrometers for MASC and VISSS and hundreds of micrometers for 2DVD and PIP), or instrumental structure (compact for MASC and 2DVD and open-field for PIP and VISSS). These sensors allow for direct microphysical investigations based on pictures of individual hydrometeors. As mentioned above, indirect retrievals of the behavior of large populations of particles within the sampling volume of remote-sensing instruments are also commonly employed, mostly by means of dual-polarization technologies \cite{Matrosov_JAS_2005}.

@@

% ---------------
 \begin{table}
 \caption{Non-exhaustive list of research works that provide assumptions or measurements of the distribution of orientation angles of ice phase particles. $\overline{\theta}$ indicates mean orientation, $\sigma$ is the standard deviation,  $\theta_{M}$ indicates the mode of the distribution, $\theta_{Q50}$ is the median, $MAD(\theta)$ is the median absolute deviation.}
 \label{table:orientations}
 \centering
 \begin{tabular}{l p{90mm}} 
 \hline
  Ref.   & Details \\
 \hline
   \cite{Matrosov_JAM_2001} & $\overline{\theta} = 0^\circ, \sigma = 3^\circ$ \\
   \hline
   \cite{Matrosov_JAS_2005} & $\overline{\theta} = 0^\circ, \sigma = 9^\circ \pm3^\circ$ (dendrites) \\
   \hline
    \cite{Kennedy_JAMC_2011}  &   $\overline{\theta} = 0^\circ, \sigma = 15^\circ$ (dendrites), 30$^\circ$ (aggregates) \\
   \hline
    \cite{Melnikov_JAOT_2013} & $\overline{\theta} = 0^\circ, \sigma  \in [2,20]^\circ$ (ice-phase clouds) \\
   \hline
    \cite{Ryzhkov_JAMC_2011} & $\overline{\theta} = 0^\circ, \sigma = 40^\circ$ (aggregates or graupel) \\
   \hline
   \cite{Hogan_JAMC_2012} & Assumed horizontal  \\
   \hline   
   \cite{Putnam_MWR_2017} & $\overline{\theta} = 0^\circ , \sigma = 20^\circ$ (snow) \newline 
   $\overline{\theta} = 0^\circ , \sigma \in [0,60]^\circ$ (graupel) \\
   \hline
   \cite{Bukovic_JAMC_2018} &  $\overline{\theta} = 0^\circ, \sigma = 10^\circ$ (dendrites) , 40$^\circ$ (aggregates) \\
   \hline
    \cite{Matsui_JGRA_2019}  & $\overline{\theta} = 0^\circ , \sigma = 20^\circ$ (aggregates) \newline 
   $\overline{\theta} = 20^\circ , \sigma = 42^\circ$ (graupel) \\
   \hline
   \cite{Garrett_GRL_2015}$^*$ & $\theta_{Q50} = 39^\circ, \theta_{M} = 13^\circ$ (aggregates) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 16^\circ$ (rimed particles) \newline
   $\theta_{Q50} = 35^\circ, \theta_{M} = 20^\circ$ (graupel) \newline
   $\theta_{Q50} = 28^\circ, \theta_{M} = 13^\circ$ (low-turbulence, all types) \newline
   $\theta_{Q50}= 40^\circ, \theta_{M} = 23^\circ$ (high-turbulence, all types) \\
   \hline
   \cite{Gergely_JGRA_2016}$^*$ &  $\theta_{Q50} \in [38,47.3]^\circ$, $MAD(\theta) \in [43,51]^\circ$ \\
   \hline
   \cite{Stout_ACP_2024}$^*$ &  $\overline{\theta} \in [0,28]^\circ$ (planar crystals, varying Re number)\\
   \hline
   \cite{Jiang_JAS_2019}$^*$ & $\theta_{M}\approx 10^\circ$ \\
   \hline
   \cite{Fitch_AMT_2021}$^*$ & $\theta_{M} = 57^\circ$ (wind $> 5$ ms$^{-1}$) \newline
   $\theta_{M} = 28^\circ$ (wind $< 1.5$ ms$^{-1}$) \\
   \hline
   \cite{Fitch_JGR_2022}$^*$ & $\theta_{M} = 3^\circ$ or  13$^\circ$ (different estimations)  \\
   \hline
   \cite{Schrom_JAS_2023}$^*$ & Beta distribution (varying parameters) \\
   \hline
    \multicolumn{2}{l}{$^{*}$Papers that present orientation statistics based on field observations.}
 \end{tabular}
 \end{table}

@@

When handling three 2D images of an individual 3D object, a primary challenge is to derive a single  estimate of orientation for each snowflake. Snowflakes are commonly modeled as ellipsoids or spheroids in previous MASC studies: the orientation of the major axis of the best-fitting ellipsoid in each 2D image is estimated, and the average of the three views is considered as a proxy for orientation \cite{Garrett_GRL_2015, Gergely_JGRA_2016, Fitch_AMT_2021, Fitch_JGR_2022}. The cited studies calculate the mean of the absolute value of the three estimates, an approach referred to here as \textit{Mean\_Abs}. Alternatively, other studies perform a 3D reconstruction of the ellipsoid that best fits all three views and take the orientation of its major axis as the estimate \cite{Jiang_JAS_2019}, a family of approaches named \textit{Ellipsoid\_Fit}. Examining Table~\ref{table:orientations}, the experimental studies using these estimators report the occurrence of non-zero modal orientation values between 10$^\circ$ and 20$^\circ$, challenging the common assumption of a distribution centered at 0$^\circ$. While the non-zero modal values reported for \textit{Ellipsoid\_Fit} in \cite{Jiang_JAS_2019} are based on a small sample of aggregate snowflakes, making evaluation difficult, we demonstrate that the appearance of non-zero modal values in the \textit{Mean\_Abs} approach simply results from averaging absolute values. This will offer a reconciliation between the common assumptions made on hydrometeor orientations and the contradiction of the first studies that used the MASC, a contradiction that was also noted by~\citeA{Garrett_GRL_2015} and discussed in the Conclusions of their work.
